> # **Popular Historic Figures Exploration**  \
> by Julian Hernández

> ## Database Information

> Pantheon  is a project developed by the Macro Connections group at 
The MIT Media Lab. It is a way to celebrate our acomplishments as a species by 
documentating our global heritage. You can find more about this dataset and 
project on [Kaggle](https://www.kaggle.com/mit/pantheon-project) or 
[Pantheon's Official Site](http://pantheon.media.mit.edu/methods)

> This dataset gathers information on the 11,341 biographies that have
presence in more than 25 languages in the Wikipedia (as of May 2013).
This dataset is not restricted to any cultural domain or time period, 
including all biographies that are present in more than 
25 different language editions of Wikipedia. 

> The dataset has 17 variables from which I choose 12 to analyze. \

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
library(GGally)
library(memisc)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
base <-read.csv("data/database.csv", stringsAsFactors=FALSE)
db <- subset(base,
             select = c(full_name, birth_year, sex, country, continent, 
                        occupation, industry, domain, article_languages,
                        page_views,average_views,
                        historical_popularity_index))


db$birth_year <- as.numeric(as.character(db$birth_year))
db$sex <- as.factor(db$sex)
db$country <- as.factor(db$country)
db$continent <- as.factor(db$continent)
db$occupation <- as.factor(db$occupation)
db$industry <- as.factor(db$industry)
db$domain <- as.factor(db$domain)
db <-subset(db, !is.na(db$birth_year))
db$antiquity <- 2013 - db$birth_year
db$hpi <- db$historical_popularity_index

str(db)
```

> ## Variable Description

> **full_name**: Name of the historical figure \
> **birth_year**: Birth Year of the historical figure in the Gregorian Calendar\
> **sex**: Biological Sex of the historical figure (Male / Female) \
> **country**: Country or modern day equivalent where the historical figure 
was born. \
> **continent**: Continent where the historical figure was born. \
> **occupation**: Profession the historical figure had. \
> **industry**: Industry of the historical figure. \
> **Domain**: Domain of Knowledge where the historical figure excelled. \
> **article_languages**: Number of Languages the historical figure biography is
present in Wikipedia. \
> **page_views**: Total number of pageviews in all languages. \
> **average_views**: Total Number of pageviews divided by number of languages the
biography is in.\
> **HPI** or **historical_popularity_index**: A more complex way to measure the 
historical impact by using the number of languagues, page views, age of the
character, and the variation in pageviews per language.


```{r echo=FALSE, message=FALSE, warning=FALSE }
summary(db)
```

# Preliminar Questions
> Is there any guiding principle for historical popularity, maybe location,
profession or time period? Which professions are more historically memorable or
significant? Is there any difference in the behavior of men and woman in the data? 
How does location influences historical popularity? Is there any difference in the 
occupations that are remembered per location?

# Univariate Plots Section

> ## Year of Birth

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}

p1 <- ggplot(db, aes(x=birth_year))+
  geom_histogram(binwidth = 50)+ 
  scale_x_continuous(breaks = seq(-5000,2000,500))+
  ggtitle("Year of Birth", subtitle = "Complete dataset")+ 
  labs(colour = "Cylinders", x = "Years", y = "# of People")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))


p2 <- ggplot(db, aes(x=birth_year))+
  geom_histogram(binwidth = 10)+ 
  ggtitle("Year of Birth", subtitle = "5% Quantile and Up")+ 
  labs( x = "Years", y = "# of People") +
  xlim( 
    0,
    2010)

grid.arrange(p1,p2)

```

> **Year of Birth**: Most globally notable figures were born on recent times.
Nevertheless there is huge variability, dates range from -3500 BC to 2005 AC.

> The data follows an upward trend towards recent times but it has a dip on the
most recent years (1995 - 2010). Which could be due to the fact that people born on those years
are fairly young and still need more time to develop their careers and skills. It 
could be also due to the specialization of skills needed to work on contemporary times, which
makes teams more necessary and teams do not appear on the dataset.

```{r echo=FALSE, message=FALSE, warning=FALSE }

ggplot(db, aes(x=birth_year))+
  geom_histogram(binwidth = 5)+ 
  scale_x_continuous(breaks = seq(1800,2010,25)) +
  ggtitle("Year of Birth", subtitle = "Births from 1800 to 2005")+ 
  labs(colour = "Cylinders", x = "Years", y = "# of People")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2))) + 
  xlim(1800, 2010)

```

> ## Sex: 

```{r echo=FALSE, message=FALSE, warning=FALSE, sex}

ggplot(db, aes(x= sex))+
  geom_bar(aes(fill = sex))+ 
  ggtitle("# of People By Sex", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))

```

> There is a huge difference in the amount of men and women in the dataset. In 
this exploration it wont be possible to explain this imbalance, due to the fact 
that we are not looking at the complete dataset of biographies in wikipedia but 
a subset, the ones translated to more than 25 
languages.  

> ## Location

```{r fig.height=25, echo=FALSE, message=FALSE, warning=FALSE }





ggplot(subset(db, continent != ""), 
       aes(x=reorder(country , country,
                     function(x)-length(x))))+
  geom_bar()+ 
  ggtitle("# of People By Country", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "Country")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))+
  coord_flip()

```


```{r echo=FALSE, message=FALSE, warning=FALSE, region }

ggplot(subset(db, continent != ""), aes( x=reorder(continent , continent,
                     function(x)-length(x))))+
  geom_bar()+ 
  ggtitle("# of People By Continent", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "Countries")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))

```

> In the past graphs we can see the distribution of great people among 
continents and countries.

> In terms of continents, most notable people come from Europe, followed by 
North America and then Asia. In terms of countries, the United States is 
followed by several european countries ( France, Germany, the UK and Italy). 


>This reflects Wikipedia bias, a bias the authors of the dataset identify as an 
issue. Wikipedia is more popular in Western countries where European history and
figures are studied and remembered. Continents with 
history that spans millenia and that have huge populations such as Asia or Africa are 
lacking in representation.

> ## Industry & Domain

```{r fig.height=15, echo=FALSE, message=FALSE, warning=FALSE}


ggplot(subset(db, continent != ""), 
       aes(x=reorder(occupation , occupation,
                     function(x)-length(x))))+
  geom_bar()+ 
  ggtitle("# of People By Occupation", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "Occupation")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))+
  coord_flip()

```


```{r echo=FALSE, message=FALSE, warning=FALSE }

ggplot(db, aes(x=reorder(industry , industry,
                     function(x)-length(x))))+
  geom_bar()+ 
  ggtitle("# of People By Industry", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "Industry")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))+
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0.3))


```
```{r echo=FALSE, message=FALSE, warning=FALSE }

ggplot(db, aes(x=reorder(domain , domain,
                     function(x)-length(x))))+
  geom_bar()+ 
  ggtitle("# of People By Domain", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "Domain")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))+
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0.3))

```

> Here we have the profession of the notable people on the dataset in decreasing
levels of granularity. We can see their field divided by: occupation, industry 
and domain.

> The most common occupation in the dataset is that of politician, followed by
several profession in the entertainment industry, football player, actor and 
writer. This translates to the distribution of great people by industry and 
domain, where most of them are located in goverment, the Arts and sports. 

> Some small bits that surprised me is the amount of religious figures that are
well know and remembered. I also though that a career on the entertainment industry
would be the leading occupation and that entertainment would be the leading domain.
Since Wikipedia is volunteer run and celebrities tend to be way more popular than
any politician. Although politician do have an historical advantage, politics as
an occupation has existed for a long time.

> ## Number of Languages

```{r echo=FALSE, message=FALSE, warning=FALSE }

p1 <- ggplot(db, aes(x = article_languages))+
  geom_histogram( binwidth = 2 )+ 
  ggtitle("# of People By Article Languages", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "# of Languages")

p2 <- ggplot(db, aes(x = article_languages))+
  geom_histogram( binwidth = 1 )+ 
  ggtitle("# of People By Article Languages", 
          subtitle = "90% of the data") + 
  labs(y = "# of People", x = "# of Languages") + 
  xlim(24, quantile(db$article_languages, probs = 0.9))

grid.arrange(p1,p2)

```

```{r echo=FALSE, message=FALSE, warning=FALSE }

summary(db$article_languages)

```


> The number of languages is skewed to the left with more than 75% of the dataset
being translated to 46 languages or less. There are several outliers, like, the biggest
value in the dataset, 214 languages, that belongs to the article of Jesus Christ, who is 
followed by Barack Obama (200 languages) and US actor Corbin Bleu (193).

> We can clearly see a downward exponential trend with languages.

> ## Total Views & Average Views

```{r echo=FALSE, message=FALSE, warning=FALSE}

p1 <- ggplot(db, aes(x = page_views))+
  geom_histogram( )+ 
  ggtitle("# of Pageviews per Biography", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "# of page views")

p2 <- ggplot(db, aes(x = page_views))+
  geom_histogram( )+ 
  ggtitle("# of Pageviews per Biography", 
          subtitle = "Log10 of Page Views per Biography") + 
  labs(y = "# of People", x = "# of page views - Log10") + 
  scale_x_log10()

grid.arrange(p1, p2)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

summary(db$page_views)

```


> Page views are higly skewed to the right, using log10 its possible to see a 
better representation of the data how it peaks around 1.000.000 views and
start declining again at 10.000.000. We can see that happening using the median,
50% of the values are below 1.000.000 and 75% of the values are below 4.000.000.

> Nevertheless page views is filled with outliers, which are people that have 
changed the course of history or impacted their fields with a breath of fresh 
ideas. The poeple with most views on Wikipedia, on this dataset are: Michael 
Jackson, Adolf Hitler and Justin Bieber. An unlikely trio.

> ## Average Views

>Average Views is obtained by dividing the total number of page views by the 
amount of languages the article is available.

```{r echo=FALSE, message=FALSE, warning=FALSE}
p1 <- ggplot(db, aes(x = average_views))+
  geom_histogram( binwidth = 5000 )+ 
  ggtitle("Average # of Pageviews per Biography", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "Average # of page views") 

p2 <- ggplot(db, aes(x = average_views))+
  geom_histogram(  )+ 
  ggtitle("Average # of Pageviews per Biography", 
          subtitle = "Log10 of Average Page Views per Biography") + 
  labs(y = "# of People", x = "Average # of page views - Log10") + 
  scale_x_log10()

grid.arrange(p1, p2)

```

> Average Page Views behaves similar to total page views, is skewed to the right
and when log10 is applied it normalizes its distribution. Nevertheless it doesnt't
measure the same as total page views. What average page views tell us is who is 
consistently more popular across several languages, but it is a mean measurement
which still means that it is susceptible to large values inflating it. Perhaps
a median would have been more useful to know the distribution of the page views.

> The biographies with the highest Average Page Views count are Kim Kardashian,
Lil Wayne and Eminem. Which could either mean that Kim has a large international 
audience or that few specific countries drive their views up, so that they have
a high page view number in comparison to the amount of languages their bio is
translated to.

> ## Historical Popularity Index (HPI)

> The HPI is calculated taking into account several key indicators. Such as the
time passed since the historical figure lived, number of times their biography
has been visited,the number of views in different languages and other key factors.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(db, aes(x = historical_popularity_index))+
  geom_histogram(binwidth = 0.5)

```

```{r echo=FALSE, message=FALSE, warning= FALSE }
summary(db$historical_popularity_index)

```


> We can see it's distribution resembles a normal one, but it a little skewed to 
the right. With most values being in the 20 - 24 range. 

========================================================

# Univariate Analysis

### What is the structure of your dataset?

> There are 11337 notable people in the dataset with 12 features 
(full_name, birth_year, sex, country, continent, occupation, industry, domain, 
article_languages, page_views, average_views, historical_popularity_index). 
There are several factors: continent, sex, industry and domain. But none of
them are ordered.

> Other notable details of the variables are: \
- A huge disparity between the quantity of men and woman. \
- Most notable people come from Europe and worked in goverment institutions. \
- The most common occupation is politician \
- The country with the most great people is the United States. \
- Most people on the dataset have between 628928 and 4485693 views. \

### What is/are the main feature(s) of interest in your dataset?

> The features in the data set I'd like to explore are gender and the
historical_popularity_index. I'd like to determine the relationship between these
variables with the rest the dataset. I'd also like to know if there are variables
that predict "historical relevance" and how good of a predictors are they.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

> I think domain, birth year, continent and the amount of languages the biography
is translated to; might give an insight to the kind of historical figure that 
becomes popular on Wikipedia. 

### Did you create any new variables from existing variables in the dataset?

> No, I only used existing variables.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this? 

> Most of the numeric variables are skewed to either the right or the left, most 
of them have lots of extreme values. In some of the variables, like page_views 
and average_views, I used log10 to transform the data so its distribution could 
be more easily seen and it could be more easily managed. 

> I also transformed birth_years from a factor variable to a numeric variable.
It was what made the most sense in order to visualize and analyze it.

> I also dropped variables such as longitude, latitude. The information they 
gave is pretty specific and is not something that a variable like continent or
country wouldn't tell me. I also dropped the state variable due to it's 
specificity, it only applied to the US.

========================================================

# Bivariate Plots Section

> There are two variables in the data I want to explore, Sex and HPI. During 
this exploration I will search for relationships between the other variables 
and those elements. 

> ## Sex:


```{r echo=FALSE, warning=FALSE, Bivariate_Plots}

p1 <- ggplot(db, aes(x = sex, y = page_views)) + 
  geom_boxplot()+
  ylim(0, quantile(db$page_views, probs = 0.9)) +
  ggtitle("Total views", subtitle = "Up to the 0.9 quantile")+ 
  labs(colour = "Cylinders", x = "Sex", y = "Page Views")

p2 <- ggplot(db, aes(x = sex, y = average_views)) + 
  geom_boxplot()+
  ylim(0, quantile(db$average_views, probs = 0.9)) +
  ggtitle("Avg Views", subtitle = "Up to the 0.9 quantile")+ 
  labs(colour = "Cylinders", x = "Sex", y = "Average Views")

p4 <- ggplot(db, aes(x = sex, y = article_languages)) + 
  geom_boxplot()+
  ggtitle("Languages", subtitle = "Up to the 0.95 quantile")+ 
  labs(colour = "Cylinders", x = "Sex", y = "Average Views")+
  ylim(0, quantile(db$article_languages, probs = 0.95)) 

p3 <- ggplot(db, aes(x = sex, y = historical_popularity_index)) + 
  geom_boxplot()+
  ggtitle("HPI", subtitle = "Whole dataset")+ 
  labs(colour = "Cylinders", x = "Sex", y = "HPI")

grid.arrange(p1,p2, p4,p3, ncol = 2, nrow = 2, top = "Sex and ...")


```

> Females tend to have more views than males on both, total views and average.
But, men tend to have more outliers on the upper part of the boxplot. The only
boxplot with a different behavior is the one for historical popularity, here men
tend to have higher values than women, but men tend to have more 
outliers, just that in this case most of the outliers are on the lower part of 
the boxplot. 

> Both Male and Females have similar distributions 
of languages in their biographies.

> ### Birth year 

```{r echo=FALSE, message=FALSE, warning=FALSE }
 p1 <- ggplot(db, aes(x=birth_year))+
  geom_histogram(binwidth = 50)+ 
  scale_x_continuous(breaks = seq(-5000,2000,1000))+
  ggtitle("Whole Dataset")+ 
  labs(colour = "Cylinders", x = "Years", y = "# of People")+
  facet_wrap(~ sex)

 p2 <- ggplot(db, aes(x=birth_year))+
  geom_histogram(binwidth = 10)+ 
  scale_x_continuous(breaks = seq(-5000,2000,500))+
  ggtitle("1000 AC - Today")+ 
  labs(colour = "Cylinders", x = "Years", y = "# of People")+
  facet_wrap(~ sex) +
  xlim(1500,2010)

 grid.arrange(p1, p2, nrow=2, top= " Notable Peope Birth Year per Sex")

```

> We see the same explosion after 1750 in both sexes, but women lag behind for 250
years before seeing a truly significant growth. More women are known in 
recent times (1750 & up) than in the rest of human history. Nevertheless men
have way more records than women, as we had seen in our previous analysis. 

> We also see the same behavior that we observed before on both genders. Both of them rise until recent times where they have a small dip.

> ### Location

```{r fig.height=25, echo=FALSE, message=FALSE }

ggplot(subset(db, continent != ""), 
       aes(x=reorder(country , country,
                     function(x)-length(x))))+
  geom_bar()+ 
  ggtitle("# of People per Gender per Country", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "Country")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))+
  coord_flip()+
  facet_wrap(~ sex)
```

```{r echo=FALSE, message=FALSE, warning=FALSE }

ggplot(subset(db, continent != ""), 
       aes(x=reorder(continent , continent,
                     function(x)-length(x))))+
  geom_bar()+ 
  ggtitle("# of People per Gender per Continent", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "Continent")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))+
  coord_flip()+
  facet_wrap(~ sex)

```

> Female historical figures were born mostly in the United States and follow a 
similar distribution to the general view we explored before. In other words, 
most of them where born in european countries or the US. One relevant difference
between the two sexes is that most of the figures with **Unknown** nationality 
are men, not women.

> On the continent scale, Europe is barely holding to the first post in terms of
women notable figures, North America is quickly catching up to them.

> The rest of the world, Africa, South America and Oceania, have almost no women
notable figures.

> ### Profession

```{r fig.height=25, echo=FALSE, message=FALSE, warning=FALSE }

ggplot(subset(db, continent != ""), 
       aes(x=reorder( occupation , occupation,
                     function(x)-length(x))))+
  geom_bar()+ 
  ggtitle("# of People per Gender per Occupation", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "Occupations")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))+
  coord_flip()+
  facet_wrap(~ sex)
```

```{r  echo=FALSE, message=FALSE, warning=FALSE }

ggplot(subset(db, continent != ""), 
       aes(x=reorder( industry , industry,
                     function(x)-length(x))))+
  geom_bar()+ 
  ggtitle("# of People per Gender per Industry", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "Industry")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))+
  coord_flip()+
  facet_wrap(~ sex)

```

```{r  echo=FALSE, message=FALSE, warning=FALSE }

ggplot(subset(db, continent != ""), 
       aes(x=reorder( domain , domain,
                     function(x)-length(x))))+
  geom_bar()+ 
  ggtitle("# of People per Gender per Domain", 
          subtitle = "Data Obtained from the Pantheon Dataset") + 
  labs(y = "# of People", x = "Domain")+
  theme_minimal()+ 
  theme_update(plot.title = element_text(size = rel(2)))+
  coord_flip()+
  facet_wrap(~ sex)

```

> At first glance we can see that among the less populated occupations in the
dataset the division between genders gave us a clear divide. Some of the 
occupations in where this divide is evident are: gymnast, pornographic actor, 
model or companion tend to be roles fulfilled by women, there are few notable 
men on these roles. Meanwhile men dominate in roles such as historian, inventor,
explorer or composer. Popular occupation with few notable women. 

> Most notable women work in the Arts, specifically as actresess. Which we can 
see reflected on the distribution on the industry and domain fields. 

> Here we see a big departure from the distribution we saw on the Domain univariate
analysis. The distribution of domains for women are remarkably different than the
one from men or the general one.

> # Bivariate Analysis: Gender Exploration

> ## Talk about some of the relationships you observed in this 
part of the investigation. How did the feature(s) of interest
vary with other features in the dataset?

> First of all, there were some differences 
in occupations between men and women. Most notable men worked in institutions 
while women worked in the Arts, especifically as actresses. The most
popular profession for men was politician.

> In terms of location, most notable women came from the Unites States.
With several European countries at it's tail. Looking at the data
from a continent perspective we see that Europe is the place where
most notable women are born, followed by the US, Asia and a lack of 
figures coming from the rest of the world. Men location distribution
is similar to the general one.

> Also, A big difference comes in terms of popularity,
women tend to have more pageviews,in total and in average, than men. 
But men tend to be have a higher HPI than women.
Both have similar distribution of languages.

> Through this exploration we could see the most common country for notable women
to come from and which field would a woman most commonly work in. Nevertheless that
is only the beginning. I would like to know from if there are relationships 
between views, country and occupation.
For example: \
- Is a Male French artist more popular than an American one? 
How about a Female one? \
- Which is the most common occupation for notable women per country? How about 
by continent? And for men? \
- What kind of occupation receives the most page views on Wikipedia?

========================================================

> # HPI

> ## Page Views & HPI

```{r echo=FALSE, warning=FALSE, message=FALSE }

ggplot(db, aes(x = hpi, y = page_views)) + 
  geom_point(alpha = 0.2) +
  ylim(0, quantile(db$page_views, probs = 0.95)) + 
  ggtitle("Initial Test Between Page Views & HPI", 
          subtitle = "Page Views & Historical Popularity Index") + 
  labs(y = "Page Views", x = "Historical Popularity Index")


```


```{r echo=FALSE, message=FALSE, warning=FALSE }

cor.test(db$hpi, db$page_views)

```

> An initial analysis of the relationship between page views and HPI, seems to 
point to a lack of relationship between the two, The scatterplot and the pearson
correlation index also guide us in that direction.

> In order to explore the relation of the page view variable with the HPI a bit 
more I decided to use log10 to normalize the distribution of page view.

```{r warning=FALSE, echo=FALSE, message=FALSE }

ggplot(db, aes(x = hpi, y = log10(page_views) )) + 
  geom_point(alpha = 0.2) +
  ylim(min(log10(db$page_views)), quantile(log10(db$page_views), probs = 0.95))+ 
  ggtitle("Log10 Test Between Page Views & HPI", 
          subtitle = "Log10(Page Views) & Historical Popularity Index") + 
  labs(y = "Log10(Page Views)", x = "Historical Popularity Index")


```

```{r echo=FALSE, message=FALSE, warning=FALSE }

cor.test(db$hpi, log10(db$page_views) )


```

> After applying log10 to the variable we see that there is still 
no clear relationship between a HPI and page views.

> I double checked using the pearson correlation index, which is lower than 
it was before, almost 0, meaning "No relation".

> ## Average Views

```{r echo=FALSE, message=FALSE, warning=FALSE }

ggplot(db, aes(x = hpi, y = average_views)) + 
  geom_point(alpha = 0.2) +
  ylim(0, quantile(db$average_views, probs = 0.95)) + 
  ggtitle("Test Between Avg Page Views & HPI", 
          subtitle = "Mean Page Views & Historical Popularity Index") + 
  labs(y = "Mean Page Views", x = "Historical Popularity Index")


```

```{r echo=FALSE, message=FALSE, warning= FALSE }

cor.test(db$hpi, db$average_views)

```

> I decided to run a similar process in average views as I ran in page_views.
On the scatterplot we can se a lack of relationship between the variables. An
hypothesis that is confirmed by the pearson correlation index, -0.08, which
indicates a lack of relationship.

> I run the test again with log10, to see if normalizing the distribution of the
variable helps.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(db, aes(x = hpi, y = log10(average_views))) + 
  geom_point(alpha = 0.2) +
  ylim(min(log10(db$average_views)), 
       quantile(log10(db$average_views), probs = 0.95)) + 
  ggtitle("Log10 Test Between Avg Page Views & HPI", 
          subtitle = "Log10 Mean Page Views & Historical Popularity Index") + 
  labs(y = "Log10 Mean Page Views", x = "Historical Popularity Index")

```

```{r echo=FALSE, message=FALSE, warning= FALSE }

cor.test(db$hpi, log10(db$average_views) )
```

> Just as with page_views, using log10 makes the pearson coefficient go down.

> Both average and total page views seem to have little relevance in the 
importance of an historical figure.

> ## Birth Year

> In order to explore correlation between hpi and birth year I tranformed 
the birth year variable to avoid negative numbers using the following formula:
**antiquity = 2013 - birth_year**.

> We substract from 2013 since the dataset was updated last on May 2013.

```{r echo=FALSE, message=FALSE, warning=FALSE }

ggplot(db, aes(x = hpi, y = antiquity)) + 
  geom_jitter(alpha = 0.5)+
  scale_y_continuous( breaks = seq(-0, 5000, 500) ) + 
  ggtitle("Older Notable figures tend to have a higher HPI", 
          subtitle = "Antiquity in years & Historical Popularity Index") + 
  labs(y = "Antiquity", x = "Historical Popularity Index")

```


> Using the scatterplot we can see that recent great people then to have the most
variance in HPI and  older historical figures tend to have higher HPI. On the 
other side recent notable people are the only ones in the dataset with 
extremely low HPI values.

```{r echo=FALSE, message=FALSE, warning= FALSE }

cor.test(db$hpi, log10(db$antiquity))

```

> Using the pearson correlation index we can see that antiquity and HPI are 
related. It could be one of the potential variables used to predict HPI.

> ## Average Languages

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(db, aes(x = hpi, y = article_languages)) + 
  geom_jitter( alpha = 0.2) +
  ylim(25, quantile(db$article_languages, probs = 0.95))+ 
  ggtitle("There is a vague positive relationship", 
          subtitle = "# of Languages & Historical Popularity Index") + 
  labs(y = "# of Languages", x = "Historical Popularity Index")

```

> There seems to be a vague positive relationship between HPI and # of Languages.
The scatterplot is too disperse. I used pearson to check the relationship 
between the two. 

```{r echo=FALSE, message=FALSE, warning=FALSE }

cor.test(db$hpi, db$article_languages)

```

> Languages and HPI have a high correlation index in comparison to page views or
average views. Nevertheless 0.4 is still too low to consider a strong relationship 
between the two.

> ## Profession & Domain

```{r fig.height=30, message= FALSE, echo=FALSE, warning=FALSE }

ggplot(db, aes(x=reorder( occupation , occupation,
                          function(x)-length(x)),
               y = hpi))+
  geom_boxplot()+
  geom_hline( yintercept= median(db$hpi), linetype="dashed", color = "red") +
  coord_flip()+ 
  ggtitle("Which occupations have a higher HPI?", 
          subtitle = "Occupation & Historical Popularity Index") + 
  labs(y = "Occupations", x = "Historical Popularity Index")
  

```


> In professions there are winners there losers. In the past graphic we can see 
are occupations that never rise above the median HPI, the worst case is the one
of swimmers which fall behind every other profession. On the other side, 
The profession with the highest median HPI is philosophy.

> Other low performing occupations include: gymnast, skater, skier or tennis 
player. 

> Other high performing occupations are: pirate, public worker or explorer.

```{r fig.height=15, message= FALSE, echo=FALSE, warning=FALSE }

ggplot(db, aes(x=reorder( industry , industry,
                          function(x)-length(x)),
               y = hpi))+
  geom_boxplot()+
  geom_hline( yintercept= median(db$hpi), linetype="dashed", color = "red") +
  coord_flip()+ 
  ggtitle("Which Industry has a higher HPI?", 
          subtitle = "Industry & Historical Popularity Index") + 
  labs(y = "Industry", x = "Historical Popularity Index")

```

```{r message= FALSE, echo=FALSE, warning=FALSE }

ggplot(db, aes(x=reorder( domain , domain,
                          function(x)-length(x)),
               y = hpi))+
  geom_boxplot()+
  geom_hline( yintercept= median(db$hpi), linetype="dashed", color = "red") +
  coord_flip()+ 
  ggtitle("Which Domain has a higher HPI?", 
          subtitle = "Domain & Historical Popularity Index") + 
  labs(y = "Domain", x = "Historical Popularity Index")


```


> In the past graphs we can see the picture painted in the occupation boxplots 
come to life.

> Industries related with sports do poorly on the HPI. Meanwhile, philosophy,
history and fine arts are the ones with the highes HPI. 

> I have to point out that even though Fine Arts is doing great other fields of 
the arts are not that lucky: Music, Film & Theater and Media Personality are
some of the worst performing industries.

> Finally it all comes together on domain, the Humanities (where philosophy is
located) soar along with Institutions. While the Arts and Sports are the fields
with the lowest HPI.

> ## Location

```{r fig.height=40, message=FALSE, warning=FALSE, echo=FALSE }

ggplot(db, aes(x=reorder( country , country,
                          function(x)-length(x)),
               y = hpi))+
  geom_boxplot()+
  geom_hline( yintercept= median(db$hpi), linetype="dashed", color = "red") +
  coord_flip()+ 
  ggtitle("Which Country has a higher HPI?", 
          subtitle = "Country & Historical Popularity Index") + 
  labs(y = "Country", x = "Historical Popularity Index")


```

```{r message=FALSE, warning=FALSE, echo=FALSE }

ggplot(subset(db, continent!=""), aes(x=reorder( continent , continent,
                          function(x)-length(x)),
               y = hpi))+
  geom_boxplot()+
  geom_hline( yintercept= median(db$hpi), linetype="dashed", color = "red") +
  coord_flip()+ 
  ggtitle("Which Continet has a higher HPI?", 
          subtitle = "Country & Historical Popularity Index") + 
  labs(y = "Country", x = "Historical Popularity Index")
```

> Most countries fall under the median HPI line. With the country with the 
lowest HPI being Swaziland. The countriy with the highest HPI shouldn't come as
a surprise since their historical colaborations are well know. It is also 
known as the birthplace of Western Civilization, Greece.

> It also important to point out the behavior of the countries with the most 
notable people, the US, UK and France. The UK and the US have their respective 
medians below the global one and have lots of outliers in either side. Meanwhile,
France behaves similarly to Germany and Italy. They have HPI higher than the global 
but also a lot of outliers pulling them down. 

> Finally I have to point out that **Unknown** country has one of the highest 
HPIs.

> In terms of domain we see the past behavior reinforced. Most continents have 
low HPI levels only Europe, Asia and Unknown have median higher than the global 
median.

> I would also like to point out that North America has most of its values below
the mean.

> ## Sex

> We analyzed the relationship between sex and HPI before. Men tend to have
higher HPI values than women.

```{r echo=FALSE, warning=FALSE, warning=FALSE}

by(db$hpi, db$sex, summary)

```

> ## General

```{r fig.height=12, echo=FALSE, warning=FALSE, message=FALSE}
set.seed(4685)
db_subset <- subset(db, select = 
                      c(antiquity, domain, continent,article_languages,
                        sex, page_views, average_views, hpi))

names(db_subset)
ggpairs(db_subset[sample.int(nrow(db_subset), 1000),])


```


# Bivariate Analysis: HPI

## Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in
the dataset?

> They main takeaway of the exploration of HPI was the lack of 
relationship between it and most variables. The exception being,
Antiquity that has a strong correlation with HPI, and # of Languages.

> In terms of categorical variables of factors. In the next phase of the
exploration I will only use the less granular format, domain for occupations and
continent for countries. 

> The Continent variable has a couple of surprises, Unknown is the highest 
median on the dataset followed by Europe or Asia. Surprisingly the US lacks 
behind. Continents like South America, Africa and Oceania have the lowest median
overall.

> In terms of profession we can see that most domain variables have a median that 
is similar to the global one. With the exception of Sports that is way below and 
Humanities & Institutions that are above the global median.

> In this categorical variables we saw that the HPI distribution between factors
is not equal, some have higher or lower HPIs. This proves helpful to our
modelling ambitions.


## Did you observe any interesting relationships between \
the other features (not the main feature(s) of interest)?

> Yes, average_views and page_views have a high correlation index. Also Number of
Language a biography has been translated to is related to number of page views it 
receives.

>Also there is a curious relation between views, antiquity and hpi. 
Just as with HPI, recent notable figures tend to have more variance
but as a figure gets older they tendto receive less views.
Which is the opossite of what happens with HPI, as a 
figure gets older it becomes more historically relevant.

========================================================

# Multivariate Plots Section

> ## Sex:

> I come to this part of the analysis with questions already in mind, like: \
- Which profession **domain** recieve more visits depending on gender? \
- Which continent has more notable women? 

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

> ## Sex, Occupation and Popularity.

```{r fig.height=12,echo=FALSE,warning=FALSE, message=FALSE }

ggplot(db, aes(x = industry, y = page_views)) + 
  geom_boxplot()+
  geom_hline( yintercept= median(db$page_views), linetype="dashed", color = "red") +
  ylim(0, quantile(db$page_views, probs = 0.9 ))+
  facet_wrap(~ sex)+
  coord_flip()
  
```

```{r fig.height=4,echo=FALSE,warning=FALSE, message=FALSE }

ggplot(db, aes(x = domain, y = page_views)) + 
  geom_boxplot()+
  geom_hline( yintercept= median(db$page_views), linetype="dashed", color = "red") +
  ylim(0, quantile(db$page_views, probs = 0.9 ))+
  facet_wrap(~ sex)+
  coord_flip()

```


> As we had seen from our analysis before, Women tend to have more views 
than men. Now we see where does views concentrate.

> Women receive more significantly more pageviews in the following 
occupations: Outlaws, Dance or Design. These are fields where women
are more popular than men in most cases. Nevertheless, Women receive
more pagevies than their male counterparts in almost every industry but
men tend to have more outliers after the 75% quantile. Which means,
women as a group tend to
be more popular, but the notable people with the most page views
in an specific industry with the most page views is likely to be male.

> We see a similar behavior on the domain variable. Women tend have more
page views than men in most domains except in Sports and institutions. But
We see the same behavior with outliers that we saw before. Although men 
as a group tend to be less popular than women as a group, men notable 
figures tend to be more viewed. The exception to this happens on
the Arts and Exploration domain, where women tend to have more views and there
are few or no male outliers. 

> ## Sex, Continent and Page Views

```{r fig.height=4,echo=FALSE,warning=FALSE, message=FALSE }

ggplot(subset(db, continent!=""), aes(x = continent, y = page_views)) + 
  geom_boxplot()+
  geom_hline( yintercept= median(db$page_views), linetype="dashed", color = "red") +
  ylim(0, quantile(db$page_views, probs = 0.9 ))+
  facet_wrap(~ sex)+
  coord_flip()
```

> Here we see the same behavior that we saw on the location variables,
Women tend to be have a higher median and 75% quantile but men have
more ouliers which end up meaning that the most viewed people on each
continent are likely men. Only North America and Oceania behave differently,
they have a higher median and 75% quantile with few or no male outliers.

> ## Sex, Location, Views and Occupation

```{r fig.height=12,echo=FALSE,warning=FALSE, message=FALSE}

ggplot(subset(db, continent!=""), aes(x = continent, y = page_views)) + 
  geom_boxplot()+
  geom_hline( yintercept= median(db$page_views), linetype="dashed", color = "red") +
  ylim(0, quantile(db$page_views, probs = 0.9 ))+
  facet_grid(domain ~ sex)+
  coord_flip()


```

> Lets start by dissecting each domain by the number of page views they 
get on each continent.

> ### Arts:
> The first thing that pops up is the lack of women artist whose birth location
is unknown. Second of all, notice that Europe has the same behavior observed 
before women tend to receive more views but men have way more outliers with
high page view amounts. Other than that we see that in Asia, Africa and North
America women tend to have more slightly more views. Artist from Oceania and
South America tend to have more pageviews if they are male.

> In other words the most popular male artists will surely come from Europe and the
female artist with more pageviews will come from North America.

> ### Business & Law:
> Women sure got the short straw on this domain. Most continents don't have
notable business women and the ones that do, North America, Asia and Europe, 
receive way less page views than their male counterparts in those continents.

> ### Exploration:
> The behavior is very similar to **Business & Law** with the addition of
African Explorers.

> ### Humanities:
Perhaps the most eye catching fact of this section is the high amount of views
that South American women tend to have. Other than that we see the a similar behavior
to other fields of knowledge, women tend to have slightly more views, men have
more outliers with higher values.

> ### Institutions:
We see the same pattern here, women have more views, men have more outliers.
Nevertheless I want to point out a couple of details, the great amount of
views women in institutions get in South America and the huge disparity 
between men and women in North America.

> ### Public Figures:
> This one is pretty mixed up. Men are popular in some continents, Women
in others. There are not that many outliers and no clear structure.

> ### Science & Technology:
>Most continents have few women scientist but in those who do we see 
they have slightly more page views in general, the only continent
that has more views than expected is Africa, where women tend to have 
way more than their male counterparts. Nevertheless we do see the 
same behavior with outliers than before.

> ### Sports:
> This is the male dominated industry. Men have tend to get more 
page views and have lots of outliers in with really high values. Women
tend to have less page views than the global median.

> **Conclusion**: The behavior we saw mean women tend to be more popular
and receive more pageviews in most fields and continents but men tend to 
have the most viewed notable people in most fields and continents
with outliers that have thousands of views more than most popular women.

> I decided to explore antiquity in relation to sex to see which how the 
different occupations and continents behaved. I decided not to explore
# of Languages due the fact that we know from past analysis that it is 
consistent among genders.

> ## Antiquity, Sex, Location and Occupation

```{r fig.height=10, echo=FALSE,warning=FALSE, message=FALSE}

ggplot(subset(db, continent!=""), aes(x = continent, y = antiquity)) + 
  geom_boxplot()+
  geom_hline( yintercept= median(db$antiquity), linetype="dashed", color = "red") +
  ylim(0, quantile(db$antiquity, probs = 0.9 ))+
  facet_grid(domain ~ sex)+
  coord_flip()

```

> There are old profession in human story, the ones related to exploration, 
institutions and the humanities (philosopy, religion) are some of them. This is
Evident when you look at t. The oldest group is the one belonging to notable
men from an unknown origin that worked in institutions, their median antiquity
is around 900 years. 

> There are few differences between sexes in this scale.

> ## HPI, Sex, Location and Occupation

```{r fig.height=10, echo=FALSE,warning=FALSE, message=FALSE}

ggplot(subset(db, continent!=""), aes(x = continent, y = hpi)) + 
  geom_boxplot()+
  geom_hline( yintercept= median(db$hpi), linetype="dashed", color = "red") +
  facet_grid(domain ~ sex)+
  coord_flip()

```

> We know that men tend to have higher HPI than women, but this division allows us
to see how deep that difference goes. Most occupations in most continents have men
with a higher HPI than women. 

> The only fields and continents in which women tend to have a higher 
historical popularity index are: Asian women working in Science, African women
working in institutions, 

# Multivariate Analysis: Sex

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

> Most of the plots confirmed what we already knew about the data. 
Women tend to have more page views. Men tend to have a higher HPI value.
And most of the values in the dataset are from the last 300 years.

> Nevertheless we still found interesting insights. Women tend to 
have more page views as a group, but men tend to have more outliers with
really high values. Which means few men receive the most visits and become the
face of their field but women an "average" women tends to receive more visits
than an "average" men.

> In terms of antiquity we see that most domains tend to have low antiquity values
with the exception of Exploration, Humanities and Institutions, the oldest of 
professions. Men and Women in these domains tend to have high antiquity values in 
continents like Europe or Asia, the other continents have values that are 
concentrated on recent times. Even though we see both men and women from ancient
times be remembered, we have to point out that men have more outliers which means
that as we go further into our past we only remember men, women do not appear on
the record.

> Finally, Sex interaction with HPI. We know that men have a higher median HPI 
and that is reflected in most combinations of Continent, Domain and Sex. There is
definetly a gap between men and women specially in the domain of business & law 
and exploration. The Domain least likely to be remembered is Sports, both men 
and women have low HPI values, that are way below the median. Women, specifically,
don't have great chances of having a high HPI if they work in the Arts domain. 
Although Men on the Art field have a higher chance of having high HPI values.

> **In Conclusion,** It is no secret that women have had the short end of the 
stick throught history and through this analysis of notable people
I wanted to dig depper into that. Although in later years there have been more 
notable women than in all of history I would like to point out that most of those
women are actresses, there are few women politicians, generals, game designers, 
historians, you name it.  I would also like to point out that even though men
have a higher HPI than women this is mostly driven by antiquity, as we saw on 
our bivariate analysis, so the lack of women in history becomes a self fulfilling
prophecy: older notable people have higher HPI, men have more antique records than
women, therefore women will constantly be under the median value. Overall, this 
was a fun dataset to analyze in terms of gender, which leads us to the next question
is HPI something we can predict using the data we have? would our predictions be
 trustworthy?



========================================================

# Multivariete Plots: HPI

> ## HPI, Languages, Continents and Domain

```{r fig.height=8, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(db, aes( x = article_languages, y = hpi )) +
  geom_jitter( aes(color = domain), alpha = 0.5) +
  xlim( min(db$article_languages),
        quantile(db$article_languages, probs = 0.95))


```

```{r fig.height=8, echo=FALSE, message=FALSE, warning=FALSE }

ggplot(subset(db, continent!=""),
       aes( x = article_languages, y = hpi )) +
  geom_jitter( aes(color = continent), alpha = 0.5) +
  xlim( min(db$article_languages),
        quantile(db$article_languages, probs = 0.95))



```

> The number of languages and HPI had a weak positive correlation between them.
Nevertheless there still are things to find, by using color we can see that there 
is a clear distribution between domains. Sports is on the bottom with lower HPI
values and domains like the Humanities or Institutions tend to stay on top. 

> When analyzing the distribution of continents in the scatterplot we see that no
such relationship emerges. There is no clear distinction or areas for each color.

> I also explored the distribution of women in the scatterplot but due to the
small numbers in the dataset they get drowned out by men. Besides there is no
clear pattern to their behavior.

> ## HPI, Antiquity, Continents and Domain

```{r  fig.height=8, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(db, aes( x = antiquity, y = hpi )) +
  geom_jitter( aes(color = domain), alpha = 0.3)+
  xlim( min(db$antiquity),
        quantile(db$antiquity, probs = 0.95))


```

```{r fig.height=8, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(subset(db, continent!=""),
       aes( x = antiquity, y = hpi )) +
  geom_jitter( aes(color = continent), alpha = 0.3)+
  xlim( min(db$antiquity),
        quantile(db$antiquity, probs = 0.95))
```

> Most records on the dataset are from very recent times. Nevertheless patterns
still emerge. 

> On the Domain Visualization we see that there is a big mesh of occupations in 
recent times with very different values. But that as we tend to go further into
the past few professions, with high HPI, remain. 

> It's also clear that must domains have lots of variances, specially in recent
times but a domain that has consistently low HPI values is Sports.

> On the other hand, continent is not a good variable to include, it doesn't 
unveils any hidden pattern in the data, the variance in each continent is too big.
Finally as we go further into the past we see that there is no clear "winner" as
it happened with Domain, many continents have old notable figures.

> ## HPI, Page Views, Continents and Domain

```{r fig.height=8, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(subset(db, continent!=""),
       aes( x = page_views, y = hpi )) +
  geom_jitter( aes(color = domain), alpha = 0.3)+
  xlim( min(db$page_views),
        quantile(db$page_views, probs = 0.95))
```

```{r fig.height=8, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(subset(db, continent!=""),
       aes( x = page_views, y = hpi )) +
  geom_jitter( aes(color = continent), alpha = 0.5)+
  xlim( min(db$page_views),
        quantile(db$page_views, probs = 0.95))
```

> In terms of pageviews we see a reflection of our pasts analysis, Domain shows 
the same behavior we saw before, with people on the Humanities having high HPI 
values but sports having low HPI. We also see that having more views doesn't 
mean you'll have a high HPI in any domain.

> In terms of continents, Europe dominates. But just as before there is a lot
of variance in each factor, for example: Europe tends to have between 20 - 25 HPI
values but they have lots of notable figures with values way below that range.
There is also no relation between any continent, page views and HPI. 

> We explored page views to see if there were relationships between the 
categorical variables and HPI we had not seen, but it gave us information we 
already had.


# Multivariate Analysis: HPI

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

> During this multivariate analyisis of the HPI variable we confirmed the
way it interacted with other variables, such as its interaction with domain,
average language and birth years. We saw how HPI tends to be higher the more 
antique a notable person is and that there are some profession that tend to be
common in ancient times.

> We see a pattern form when you apply color to the language and HPI scatterplot,
with people working in the humanities and institutions having a higher HPI and
people on sports having a lower HPI.

> Finally I created visualizations using page_views to explore possible 
relationship that might have passed unnoticed before. But no such thing happened.

> Although we found several relations in the dataset there are few that are
highly correlated with the HPI variable. Which could prove tough when building
the model.


### Were there any interesting or surprising interactions between features?
> I created several test visualizations for sex, hpi and other variables like:
average_language, antiquity or page_views. It is obvious looking at the imbalance
between the number of men and women but I was surprised by how few women are
in the dataset.

> ## General 

> A way to look at relationship between variables before creating a model.

```{r fig.height=12, echo=FALSE, warning=FALSE, message=FALSE}
set.seed(4685)
db_subset <- subset(db, select = 
                      c(antiquity, domain, continent,article_languages,
                        sex, page_views, average_views, hpi))

names(db_subset)
ggpairs(db_subset[sample.int(nrow(db_subset), 1000),])
```

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

> Through the analysis I found variables I had to include in the model:
- **Birth year** has a high correlation with HPI and we know that older figures tend 
to have higher HPI. I would like to include it. \
- **# of Languages**, give an idea of how much of a global present does the notable
figure has which is linked to the HPI. \
- **Sex**: Men tend to have a higher HPI than women, that fact could help the model 
predict outcomes. \
- **Domain**: We have seen that some domains have higher HPI than other. \


```{r echo=FALSE, message=FALSE, warning=FALSE}
  m1 <- lm(I(hpi) ~ I(antiquity), data = db)
  m2 <- update(m1, ~ . + domain)
  m3 <- update(m2, ~ . + article_languages)
  m4 <- update(m3, ~ . + average_views)
  m5 <- update(m4, ~ . + sex)
  mtable(m1, m2, m3, m4, m5, sdigits = 3)
```

```{r echo=FALSE,message=FALSE, warning=FALSE }
summary(m5)
```


> Overall, the model is not very trustworthy. It R and R-Squared values are very
low which means that the model fails to explain most of the variance around its 
mean. But the p-value or significance code of most variables is way below 0.05 
which means the variables have a significant relationship with HPI.

> The model could use some improvement but with the variables at our disposal
is the most accuarate one we could have built.

------------------------------------------------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

ggplot(db, aes(x=hpi))+
  geom_histogram(binwidth = 0.33)+ 
  scale_x_continuous(breaks = seq(10,40,1)) +
  ggtitle("Distribution of HPI Values", subtitle = "Complete Dataset")+ 
  labs(colour = "Cylinders", x = "HPI", y = "# of People")+
  theme_light() + 
  theme_update(plot.title = element_text(size = rel(2)))

```

### Description One
> One of the guiding variables in our analysis. HPI is a little skewed to the 
right but doesn't have a long tail. Most notable people have HPI values around
22 and 26, the max values is 33. Even among notable people in history there 
are some more notable than others.


### Plot Two
```{r echo=FALSE, warning = FALSE, Plot_Two}


library(wesanderson)

ggplot(db, aes( x = antiquity, y = hpi )) +
  geom_point( shape = 21, aes(color = domain), 
              alpha = 0.8, size = 0.8)+
  xlim( min(db$antiquity),
        quantile(db$antiquity, probs = 0.99))+
  ggtitle("Ancient Notable people tend to have higher HPI",
          subtitle = "HPI and Antiquity colored by Domain")+ 
  labs(colour = "Domain", y = "HPI (Historical Popularity Index)",
       x = "Years")+
  theme_light() + 
  theme_update(plot.title = element_text(size = rel(2))) +
  theme_update(legend.position = "bottom")

```

### Description Two
> This plot show us some of the variables that according to pearson correlation
are related to HPI. We see that the more time has passed since a notable person
birth, the higher their HPI value will be. It also shows us that the further you
go into the past the higher the probability that you belong to certain industries.
It also shows us how some industries have really lows HPI values, like sports. 

> Overall this graph gives us many key insights about the dataset in a compact 
way.

### Plot Three
```{r fig.height=10, echo=FALSE, warning=FALSE, Plot_Three}

women <- ggplot(db, aes( x = sex))+
  geom_bar( aes( fill = sex))+
  ggtitle(" # of People", subtitle = "Complete Dataset")+ 
  labs(colour = "Cylinders", x = "Sex", y = "# of People")+
  theme_update(legend.position = "none")

p1 <- ggplot(db, aes(x = sex, y = page_views)) + 
  geom_boxplot( aes( color = sex) )+
  ylim(0, quantile(db$page_views, probs = 0.9)) +
  ggtitle("Total views", subtitle = "Up to the 0.9 quantile")+ 
  labs(colour = "Cylinders", x = "Sex", y = "Page Views")+
  theme_update(legend.position = "none")


p2 <- ggplot(db, aes(x = sex, y = average_views)) + 
  geom_boxplot(aes( color = sex))+
  ylim(0, quantile(db$average_views, probs = 0.9)) +
  ggtitle("Avg Views", subtitle = "Up to the 0.9 quantile")+ 
  labs(colour = "Cylinders", x = "Sex", y = "Average Views")

p4 <- ggplot(db, aes(x = sex, y = article_languages)) + 
  geom_boxplot(aes( color = sex))+
  ggtitle("Languages", subtitle = "Up to the 0.95 quantile")+ 
  labs(colour = "Cylinders", x = "Sex", y = "# of Languages")+
  ylim(0, quantile(db$article_languages, probs = 0.95)) 

p3 <- ggplot(db, aes(x = sex, y = historical_popularity_index)) + 
  geom_boxplot(aes( color = sex))+
  ggtitle("HPI", subtitle = "Whole dataset")+ 
  labs(colour = "Cylinders", x = "Sex", y = "HPI")

age <- ggplot(db, aes(x = sex, y = antiquity)) + 
  geom_boxplot(aes( color = sex))+
  ggtitle("Antiquity", subtitle = "85% quantile")+ 
  labs(colour = "Cylinders", x = "Sex", y = "Years")+
  ylim(0, quantile(db$antiquity, probs = 0.90)) 


grid.arrange(women, p1,p2, p4,p3, age, ncol = 2, nrow = 3, top = "Sex and ...")


```

### Description Three

> I spent a lot of time analyzing how women behaved on the dataset in comparison
to men. So I would like to give an overview, the first thing that we need to 
adresss is the huge difference in number between male and female great people. 


> Nevertheless these women know their stuff and tend to have more average and 
totalviews than men, by a lot, but men tend to have more outliers. Which means
a few men tend to be the notable people with the more views, but men as a group
tend to receive less views than women.

> In terms of Languages both men and women have been translated a similar number
of times. They have an almost identical distribution. 

> Men tend to have higher HPI, meaning that more men considered more "important" 
in historical terms than women. Its important to point out that the 25 quantile
for males is higher than the median for women, so, 75% of male notable figures
have a higher HPI than 50% of women.

> Finally, most notable people in the list were born quite recently. Most women
in the dataset where born 150 years ago or less. Men, on the other hand, also 
exploded quite recently but cover a longer period of time than women. 


------

# Reflection

> The main struggle was the huge amount of categorical variables with lots of 
factors like country or occupation. The information they had was interesting and
fun to be explored but the huge amount of factors made it hard to analyze and
compare. I really liked to see how all of this notable people where distributed
among different the different categories, gender, occupation and location. There
where also some oddities here and there like actor Corbien Bleu having his bio
translated into so many languages. 

> There is still so much you can do with this dataset. The people at Pantheon 
created a Golden Age variable that lets you know when a certain country
had more notable people than other nations on the same period of time. It could
also be nice to see how the porcentage of great people by proffesion and continent
changes distribution per time period.

> Overall is a cool dataset.

